---
title: "Untitled"
author: "Ellen"
date: "1/18/2021"
output: html_document
---


# Processing data

```{r}
library("tidyverse")

```

## Process climate data

1) Download [precipitation data](http://data.ceda.ac.uk/badc/cru/data/cru_ts/cru_ts_4.04/data/pre) onto your local machine (large files). You will have to either create an account or log in. Download 4 time periods (you want the files with the 'nc' in the name):
- 1981-1990
- 1991-2000
- 2001-2010
- 2011-2019

```{r setup, include=FALSE}
sites <- read_csv('./data/NutNetGreening_2019.10.15_ee.csv')
localdir <- "/Users/ellen/Desktop/Ellen/Guelph/Project_Andrew phenology/CRU"
  
### create precipitation data for the 4 distinct time periods
cruv <- "pre"
timeframe <- '1981.1990.'
source("./get-CRU-TS.R")

timeframe <- '1991.2000.'
source("./get-CRU-TS.R")

timeframe <- '2001.2010.'
source("./get-CRU-TS.R")

timeframe <- '2011.2019.'
source("./get-CRU-TS.R")
```


2) Download [temperature data](http://data.ceda.ac.uk/badc/cru/data/cru_ts/cru_ts_4.04/data/tmp) onto your local machine (large files). You will have to either create an account or log in. Download 4 time periods (you want the files with the 'nc' in the name):
- 1981-1990
- 1991-2000
- 2001-2010
- 2011-2019
```{r temperaturedata, include = F}
cruv <- 'tmp'
timeframe <- '1981.1990.'
source("./get-CRU-TS.R")

timeframe <- '1991.2000.'
source("./get-CRU-TS.R")

timeframe <- '2001.2010.'
source("./get-CRU-TS.R")

timeframe <- '2011.2019.'
source("./get-CRU-TS.R")
```

